---
title: "Logistic Regression"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Logistic Regression in sparklyr 

- Outline of this in sparklyr
- Spark ML
- predictive vs inferential
- Adapt from [Spark ML overview](https://spark.rstudio.com/guides/mlib.html)?


This section shows how to use the `Spark ML` functions `ml_logistic_regression`, `ml_predict` and `ml_generalised_linear_regression` in `sparklyr` to generate a logistic regression model. The syntax used to call these functions is somewhat similar to logistic regression functions in R. However, there are some additional steps to take in preparing the data before the model can be run successfully. 

### Preparing the data
We will read the data in as follows:

```{r, eval = FALSE}
sc <- sparklyr::spark_connect(
  master = "local[2]",
  app_name = "sparklyr-lab2",
  config = sparklyr::spark_config())


sparklyr::spark_connection_is_open(sc)

rescue_path_parquet = "/training/rescue_clean.parquet"
rescue <- sparklyr::spark_read_parquet(sc, rescue_path_parquet)
dplyr::glimpse(rescue)
```

Let's see if we can use the `rescue` data to predict whether an animal is a cat or not.

```{r, eval = FALSE}
# Create is_cat column to contain target variable and select relevant predictors
rescue_cat <- rescue %>% 
  dplyr::mutate(is_cat = ifelse(animal_group == "Cat", 1, 0)) %>% 
  sparklyr::select(typeofincident, 
                   engine_count, 
                   job_hours, 
                   hourly_cost, 
                   total_cost, 
                   originofcall, 
                   propertycategory,
                   specialservicetypecategory,
                   incident_duration,
                   is_cat)

dplyr::glimpse(rescue_cat)
```
Examining the dataset, there are a few columns we have selected above that are of the type "character" when they should be numeric. We can convert all of these to numeric values by running the following:

```{r, eval = FALSE}
rescue_cat <- rescue_cat %>% 
  dplyr::mutate(across(c(engine_count:total_cost, incident_duration), 
                ~as.numeric(.)))

dplyr::glimpse(rescue_cat)
```

### Missing values

You may already be familiar with functions such as `glm` for performing logistic regression in R. Many of these functions are quite user-friendly and will automatically account for issues such as missing values in the data. However, the `Spark ML` functions in `sparklyr`require the user to correct for these issues before running the regression functions.

```{r, eval = FALSE}

# Get the number of NAs in the dataset by column
rescue_cat %>%
  dplyr::summarise_all(~sum(as.integer(is.na(.)))) %>% 
  print(width = Inf)

# There are 38 missing values in the 4 of the columns 
# We can see that these are all on the same rows:  
rescue_cat %>%
  sparklyr::filter(is.na(total_cost)) %>%
  print(n=38)

# For simplicity, we will just filter out these rows:
rescue_cat <- rescue_cat %>%
    sparklyr::filter(!is.na(total_cost)) 
  
# Double check we have no NAs left 
rescue_cat %>%
  dplyr::summarise_all(~sum(as.integer(is.na(.)))) %>% 
  print(width = Inf)

```

#### Running a logistic regression model

Running a logistic regression model in `sparklyr` can be done using similar syntax to R by using the `ml_logistic_regression` function:

```{r, eval = FALSE}
# Run the model
glm_out <- ml_logistic_regression(rescue_cat, 
                                 formula = "is_cat ~ typeofincident + engine_count + job_hours + hourly_cost + originofcall + propertycategory + specialservicetypecategory")
```
Model predictions can be accessed using `ml_predict`:

```{r, eval = FALSE}
# Get model predictions
ml_predict(glm_out) %>% 
  print(n = 20, width = Inf)
```

### Inferential analysis

The `Spark ML` functions available in sparklyr were largely developed with predictive analysis in mind. This means that they have been built primarily to specify a regression model and retrieve the prediction (as we have seen above), with little information in between.

When conducting analysis at ONS, we are often not interested in predicting unknown outcomes, but instead understanding the relationship between the independent variables and the probability of the outcome. This is what is referred to as *inferential analysis* in this section.

The regression coefficients from the model above can be displayed by using the `tidy` function from the `broom` package:

```{r, eval = FALSE}
# Run the model
glm_out <- ml_logistic_regression(rescue_cat, 
                                 formula = "is_cat ~ typeofincident + engine_count + job_hours + hourly_cost + originofcall + propertycategory + specialservicetypecategory")

# View coefficients
broom::tidy(glm_out)
```
This does not provide a lot of information. The `ml_logistic_regression` function only produces regression coefficients and does not produce standard errors. This means that confidence intervals cannot be calculated. 

Instead, we must use the function `ml_generalised_linear_regression`. Here, we need to specify the `family` and `link` arguments as "binomial" and "logit" respectively in order to run a logistic regression model that produces standard errors.

```{r, eval = FALSE}
# Run model with ml_generalized_linear_regression
glm_out <- ml_generalized_linear_regression(rescue_cat, 
                                 formula = "is_cat ~ engine_count + job_hours + hourly_cost + originofcall + propertycategory + specialservicetypecategory", 
                                 family = "binomial", 
                                 link = "logit")

# View a tibble of coefficients, std. error, p-values
broom::tidy(glm_out)
```
Individual statistics of interest can also be accessed directly using `ml_summary`, for example, to get the standard errors:

```{r, eval = FALSE}
c("coefficient_standard_errors") %>%
  purrr::map(~ml_summary(glm_out, .x, allow_null = TRUE)) %>%
  purrr::map_if(is.function, ~.x())
```
Other statistics can also be accessed individually in this way by replacing "coefficient_standard_errors" with the statistic of interest:

```
ml_summary(glm_out)
GeneralizedLinearRegressionTrainingSummary 
 Access the following via `$` or `ml_summary()`. 
 - aic() 
 - degrees_of_freedom() 
 - deviance() 
 - dispersion() 
 - null_deviance() 
 - num_instances() 
 - prediction_col 
 - predictions 
 - rank 
 - residual_degree_of_freedom() 
 - residual_degree_of_freedom_null() 
 - residuals() 
 - coefficient_standard_errors() 
 - num_iterations 
 - solver 
 - p_values() 
 - t_values()
```
Unfortunately, we have not yet found/built functionality in sparkylr to calculate exact confidence intervals. Instead, we can approximate the 95% confidence intervals by multiplying the standard errors by 1.96 and subract/add this to the estimate. As sparklyr regression functions are only necessary on datasets with a large number of observations, this approximation is sufficient.

```{r, eval = FALSE}
broom::tidy(glm_out) %>%
  dplyr::mutate(lower_ci = estimate - (1.96 * std.error),
                upper_ci = estimate + (1.96 * std.error))
```

#### Singularity and perfect complements

We also need to be careful to check that variables included in the model take more than one value and that they don't combine with other variables to create perfect complements. In R, using functions such as `glm`, these variables would be dropped automatically. However, this in not that case in `sparklyr`. For example, the code below runs a logistic regression model using the `ml_generalized_linear_regression` function

```{r, eval = FALSE}
glm_out <- ml_generalized_linear_regression(rescue_cat, 
                                 formula = "is_cat ~ typeofincident + engine_count + job_hours + hourly_cost + originofcall + propertycategory + specialservicetypecategory", 
                                 family = "binomial", 
                                 link = "logit")
```
In this case, running the code above will produce an error since the `typeofincident` column only contains one variable, `Special Service`. 






#### Selecting reference categories and encoding categorical variables

When including categorical variables as independent variables in a regression, one of the categories must act as the reference category. The coefficients for all other categories are relative to the reference category. 

By default, the `ml_generalized_linear_regression` function in R will select the least common category as the reference category before running the model, without input from the user. It also implicitly applies one-hot encoding to the categories so that they can be represented as a binary numerical value. 

For example, `specialservicetypecategory` has four unique values: Animal Rescue From Below Ground, Animal Rescue From Height, Animal Rescue From Water, and Other Animal Assistance. 

```{r, eval = FALSE}
rescue_cat %>% 
  dplyr::count(specialservicetypecategory) %>% 
  dplyr::arrange(n)
```

In this case, `ml_generalized_linear_regression` would select "Animal Rescue From Water" as the reference category, since there are only 343 instances of it in the data. For the purpose of running the model, it would also create a new column in the data for each of the other categories, assigning a 1 or 0 for each row depending on whether that entry belongs to that category or not. Regression coefficients would then be shown for each of these new columns in the final model, relative to the to the "Animal Rescue From Water" reference category. 

If we wanted to specify which category we want to choose as our reference category in R, this could be done quite simply using **factors**. For example, to choose "Animal Rescue From Height" as our reference category instead, we would just need to mutate the `specialservicetypecategory` as a factor and specify the reference category as follows:

```{r, eval = FALSE}
rescue_cat <- rescue_cat %>% 
  mutate(specialservicetypecategory_ref = relevel(as.factor(specialservicetypecategory), 
                  ref = "Animal Rescue From Height"))
```

This also makes it very easy to specify reference categories for multiple columns at once by adding in additional mutate terms.

Unfortunately, factors do not exist in Spark and therefore are not a valid data type in sparklyr. So we need to find an alternative way of specifying reference categories.

Instead, we can make use of the one-hot encoding concept and two of `sparklyr`'s **Feature Transformers** (`ft_string_indexer` and `ft_one_hot_encoder`) to achieve this. These functions have limited functionality, so we must first manipulate it such that the reference category will be ordered last when using `ft_string_indexer`, and then we can drop the last category using `ft_one_hot_encoder`. A convenient way of doing this is to order the categories in *descending alphabetical* order and ensuring that our chosen category will be ordered last by adding an appropriate prefix to it. For example, adding `000_`:

```{r, eval = FALSE}
rescue_cat_ohe <- rescue_cat %>%
  dplyr::mutate(region = ifelse(specialservicetypecategory == "Animal Rescue From Height",
                         "000_Animal Rescue From Height",
                         specialservicetypecategory)) %>%
  sparklyr::ft_string_indexer(input_col = "specialservicetypecategory", 
                    output_col = "specialservicetypecategory_idx",
                    string_order_type = "alphabetDesc") %>%
  sparklyr::ft_one_hot_encoder(input_cols = c("specialservicetypecategory_idx"), 
                     output_cols = c("specialservicetypecategory_ohe"), 
                     drop_last = TRUE) %>%
  sparklyr::sdf_separate_column(column = "specialservicetypecategory_ohe",
                      into = c("specialservicetypecategory_Animal Rescue From Water", 
                               "specialservicetypecategory_Animal Rescue From Below Ground", 
                               "specialservicetypecategory_Other Animal Assistance")) %>% 
  sparklyr::select(-ends_with(c("_ohe", "_idx")))
```
Where `sdf_separate_column` has been used to separate encoded variables into individual columns and the intermediate columns with "_ohe" and "_idx" suffixes have been dropped once the process is complete. This can be repeated for every categorical variable as desired. 

```{r, eval = FALSE}
# originofcall
rescue_cat_ohe <- rescue_cat_ohe %>%
  dplyr::mutate(region = ifelse(originofcall == "Person (mobile)",
                         "000_Person (mobile)",
                         originofcall )) %>%
  sparklyr::ft_string_indexer(input_col = "originofcall", 
                    output_col = "originofcall_idx",
                    string_order_type = "alphabetDesc") %>%
  sparklyr::ft_one_hot_encoder(input_cols = c("originofcall_idx"), 
                     output_cols = c("originofcall_ohe"), 
                     drop_last = TRUE) %>%
  sparklyr::sdf_separate_column(column = "originofcall_ohe",
                      into = c("originofcall_Ambulance", 
                               "originofcall_Police", 
                               "originofcall_Coastguard",
                               "originofcall_Person (land line)",
                               "originofcall_Not Known",
                               "originofcall_Person (running Call)",
                               "originofcall_Other Frs")) 
# propertycategory
rescue_cat_ohe <- rescue_cat_ohe %>%
  dplyr::mutate(region = ifelse(propertycategory == "Dwelling",
                         "000_Dwelling",
                         propertycategory)) %>%
  sparklyr::ft_string_indexer(input_col = "propertycategory", 
                    output_col = "propertycategory_idx",
                    string_order_type = "alphabetDesc") %>%
  sparklyr::ft_one_hot_encoder(input_cols = c("propertycategory_idx"), 
                     output_cols = c("propertycategory_ohe"), 
                     drop_last = TRUE) %>%
  sparklyr::sdf_separate_column(column = "propertycategory_ohe",
                      into = c("propertycategory_Outdoor", 
                               "propertycategory_Road Vehicle", 
                               "propertycategory_Non Residential",
                               "propertycategory_Boat",
                               "propertycategory_Outdoor Structure",
                               "propertycategory_Other Residential"))

# specialservicetype
rescue_cat_ohe <- rescue_cat_ohe %>%
  dplyr::mutate(region = ifelse(specialservicetype == "Animal Rescue From Height - Domestic Pet",
                         "000_Animal Rescue From Height - Domestic Pet",
                         specialservicetype)) %>%
  sparklyr::ft_string_indexer(input_col = "specialservicetype", 
                    output_col = "specialservicetype_idx",
                    string_order_type = "alphabetDesc") %>%
  sparklyr::ft_one_hot_encoder(input_cols = c("specialservicetype_idx"), 
                     output_cols = c("specialservicetype_ohe"), 
                     drop_last = TRUE) %>%
  sparklyr::sdf_separate_column(column = "specialservicetype_ohe",
                      into = c("specialservicetype_Animal Rescue From Water - Bird", 
                               "specialservicetype_Animal Assistance - Lift Heavy Domestic Animal", 
                               "specialservicetype_Animal Rescue From Water - Domestic Pet",
                               "specialservicetype_Animal Harm Involving Domestic Animal",
                               "specialservicetype_Animal Harm Involving Livestock",
                               "specialservicetype_Animal Rescue From Height - Farm Animal",
                               "specialservicetype_Animal Harm Involving Wild Animal",
                               "specialservicetype_Animal Assistance Involving Domestic Animal - Other Action",
                               "specialservicetype_Animal Rescue From Height - Bird",
                               "specialservicetype_Wild Animal Rescue From Below Ground",
                               "specialservicetype_Animal Rescue From Below Ground - Bird",
                               "specialservicetype_Assist Trapped Domestic Animal",
                               "specialservicetype_Animal Rescue From Below Ground - Domestic Pet",
                               "specialservicetype_Assist Trapped Wild Animal", 
                               "specialservicetype_Wild Animal Rescue From Height",
                               "specialservicetype_Wild Animal Rescue From Water Or Mud",
                               "specialservicetype_Animal Assistance Involving Wild Animal - Other Action",
                               "specialservicetype_Animal Rescue From Water - Farm Animal",
                               "specialservicetype_Animal Assistance Involving Livestock - Other Action",
                               "specialservicetype_Assist  Trapped Livestock Animal",
                               "specialservicetype_Animal Assistance - Lift Heavy Wild Animal",
                               "specialservicetype_Animal Rescue From Below Ground - Farm Animal",
                               "specialservicetype_Animal Assistance - Lift Heavy Livestock Animal"
                               ))

# remove _idx and _ohe intermediate columns
rescue_cat_ohe <- rescue_cat_ohe %>% 
  sparklyr::select(-ends_with(c("_ohe", "_idx")))

```

##### Functionalising

Repeating this process for every categorical variable you wish to specify the reference category for could be very time consuming. You may find it helpful to instead use or adapt 

**Unsure how to load package - check details and add to this section later **







### Inferential analysis
**Adapt from Ted's presentation materials**


The guide below shows how to find regression coefficients, test significance, calculate confidence intervals and rebase categorical variables to produce regression output tables for inferential purposes.




